04_DESeq_sexes

### 2019-12-22 analysis/04_LDA.Rmd
lda working now

```{r setup, include = FALSE}
library(tidyverse)
library(DESeq2)
library(cowplot)
library(pheatmap)
library(viridis)
library(forcats)
library(caret) # LDA analysis
library(MASS) # LDA analysis
library(ggtext) # for pretty plots
library(ctsGE)

library(BiocParallel)
register(MulticoreParam(6))

@@

## DEseq2 analysis on characterization. Looking at treatment AND sex for each tissue

@@

characterization <- colData %>% filter(study == "charcterization")
manipluation <- colData %>% filter(study == "manipulation")

vsdshyp <- read.csv("../results/06_hypallvsd.csv")
vsdspit <- read.csv("../results/06_pitallvsd.csv")
vsdsgon <- read.csv("../results/06_gonallvsd.csv")

vsdall <- full_join(vsdshyp, vsdspit)
vsdall <- full_join(vsdall, vsdsgon)
head(vsdall)
```

@@

```{r LDAdfs}
# LDA for treatment (9 groups)
LDA.hyp1 <- LDAdata.treatment(vsd.hyp, colDataHyp)
LDA.pit1 <- LDAdata.treatment(vsd.pit, colDataPit)
LDA.gon1 <- LDAdata.treatment(vsd.gon, colDataGon)

# LDA for hypothesis (3 groups)

LDA.hyp2 <- LDAdata.hypothesis(vsd.hyp, colDataHyp)
LDA.pit2 <- LDAdata.hypothesis(vsd.pit, colDataPit)
LDA.gon2 <- LDAdata.hypothesis(vsd.gon, colDataGon)
```

```{r LDA, fig.width = 9}
# treatment
a <- LDAplot.treatment(LDA.hyp1 , "Hypothalamus model 1", "parental state ~ .  0.277 pred. acc.",
                        "LD1, F = 7.1504779", "LD2, F = 2.4219585")
b <- LDAplot.treatment(LDA.pit1 ,  "Pituitary model 1", "parental state ~ .  0.405 pred. acc.",
                        "LD1, F = 9.6144160", "LD2, F = 3.7773314")
c <- LDAplot.treatment(LDA.gon1 , "Gonad model 1", "parental state ~ .  has 0.297 pred. acc.",
                        "LD1, F = 6.137500", "LD2, F = 3.320437")

c <- c + theme(legend.direction = "horizontal")
mylegend <- get_legend(c)

temp <- plot_grid(a + theme(legend.position = "none"),
                  b + theme(legend.position = "none"),
                  c + theme(legend.position = "none"), nrow = 1)

plot_grid(temp, mylegend, nrow = 2, rel_heights = c(1,0.3))

# hypothesis
d <- LDAplot.hypothesis(LDA.hyp2 , "Hypothalamus model 2", "parental stage ~ .  0.432 pred. acc.",
                        "LD1, F = 7.386458", "LD2, F = 3.720794")
e <- LDAplot.hypothesis(LDA.pit2 ,  "Pituitary model 2", "parental stage ~ .  0.473 pred. acc.",
                        "LD1, F = 7.846659", "LD2, F = 5.076861")
f <- LDAplot.hypothesis(LDA.gon2 , "Gonad model 2", "parental stage ~ .  has 0.368 pred. acc.",
                        "LD1, F = 6.029069", "LD2, F = 3.837591")

f <- f + theme(legend.direction = "horizontal")
mylegend <- get_legend(f)

temp <- plot_grid(d + theme(legend.position = "none"),
                  e + theme(legend.position = "none"),
                  f + theme(legend.position = "none"), nrow = 1)

plot_grid(temp, mylegend, nrow = 2, rel_heights = c(1,0.3))
```

>>>>

```{r setup, include = FALSE}
library(tidyverse)
library(cowplot)
library(forcats)
library(caret) # LDA analysis
library(MASS) # LDA analysis

@@

## LDA analysis of characterization and manipulation

@@

charHyp <- colData %>% filter(study == "charcterization", tissue == "hypothalamus") %>% droplevels()
charPit <- colData %>% filter(study == "charcterization", tissue == "pituitary") %>% droplevels()
charGon <- colData %>% filter(study == "charcterization", tissue == "gonad") %>% 
  mutate(tissue = fct_recode(tissue, "gonads" = "gonad")) %>% droplevels()

manipHyp <- colData %>% filter(study == "manipulation", tissue == "hypothalamus") %>% droplevels()
manipPit <- colData %>% filter(study == "manipulation", tissue == "pituitary") %>% droplevels()
manipGon <- colData %>% filter(study == "manipulation", tissue == "gonad") %>% 
  mutate(tissue = fct_recode(tissue, "gonads" = "gonad")) %>% droplevels()

selectvsd <- function(pathtofile, colData){

  df <- read.csv(pathtofile, row.names = 1)
  savecols <- as.character(colData$V1) 
  savecols <- as.vector(savecols) 
  df <- df %>% dplyr::select(one_of(savecols)) 
  df <- as.data.frame(t(df))
  df$V1 <- row.names(df)
  return(df)
}

vsd.hyp.train <- selectvsd("../results/06_hypallvsd.csv",  charHyp)
vsd.pit.train <- selectvsd("../results/06_pitallvsd.csv",  charPit)
vsd.gon.train <- selectvsd("../results/06_gonallvsd.csv",  charGon)

vsd.hyp.test <- selectvsd("../results/06_hypallvsd.csv",  manipHyp)
vsd.pit.test <- selectvsd("../results/06_pitallvsd.csv",  manipPit)
vsd.gon.test <- selectvsd("../results/06_gonallvsd.csv",  manipGon)
```

@@

```{r LDAdfs}

LDanalysis <- function(trainsamples, traindata, testdata, testsamples){

  train.data <- left_join(trainsamples, traindata) %>%
    dplyr::select(-V1, -bird, -tissue, -group, -study, -sex)

  test.data <- testdata

  # Normalize the data. Categorical variables are automatically ignored.
  # Estimate preprocessing parameters
  preproc.param <- train.data %>% 
    preProcess(method = c("center", "scale"))

  # Transform the data using the estimated parameters
  train.transformed <- preproc.param %>% predict(train.data)
  test.transformed <- preproc.param %>% predict(test.data)

  # LDA analysis
  # Fit the model
  model <- lda(treatment~ ., data = train.transformed)
  # Make predictions
  predictions <- model %>% predict(test.transformed)

  # Model accuracy
  print("model accuracy")
  print("predictions$class==test.transformed$treatment)")
  print(mean(predictions$class==test.transformed$treatment))

  # results
  print("the samples sizes")
  print(model$counts)

  print("the prior probabilities used")
  print(model$prior)

  print("svd: the singular values, which give the ratio of the between- and within-group standard deviations on the linear discriminant variables. Their squares are the canonical F-statistics.")
  print(model$svd)

  #  predictions
  predictions <- model %>% predict(test.transformed)
  head(predictions)

  # Predicted classes
  #print(predictions$class, 6)
  # Predicted probabilities of class memebership.
  #print(predictions$posterior, 6) 
  # Linear discriminants
  #print(predictions$x, 3)

  predictedstage <-  predict(model, test.transformed)$class
  testsamples$predictedstage <- predictedstage

  lda.data <- cbind(testsamples, predictions$x)

  return(lda.data)
}

LDA.hyp <- LDanalysis(charHyp, vsd.hyp.train, vsd.hyp.test, manipHyp)
LDA.pit <- LDanalysis(charPit, vsd.pit.train, vsd.pit.test, manipPit)
LDA.gon <- LDanalysis(charGon, vsd.gon.train, vsd.gon.test, manipGon)
```

```{r LDAplots}
#myshapes = c("hypothalamus" = 20,  "pituitary" = 17,  "gonads" = 15)

a <- ggplot(LDA.hyp, aes(x = LD1, LD2, color = predictedstage)) + geom_point(shape = 20) + theme(legend.position = "none") + scale_color_manual(values = colorscharmaip)
b <- ggplot(LDA.hyp, aes(x = LD1, LD2, color = treatment)) + geom_point(shape = 20) + theme(legend.position = "none") + scale_color_manual(values = colorscharmaip)

plot_grid(a,b)

a <- ggplot(LDA.pit, aes(x = LD1, LD2, color = predictedstage)) + geom_point(shape = 17) + theme(legend.position = "none") + scale_color_manual(values = colorscharmaip)
b <- ggplot(LDA.pit, aes(x = LD1, LD2, color = treatment)) + geom_point(shape = 17) + theme(legend.position = "none") + scale_color_manual(values = colorscharmaip)

plot_grid(a,b)

a <- ggplot(LDA.gon, aes(x = LD1, LD2, color = predictedstage)) + geom_point(shape = 15) + theme(legend.position = "none") + scale_color_manual(values = colorscharmaip)
b <- ggplot(LDA.gon, aes(x = LD1, LD2, color = treatment)) + geom_point(shape = 15) + theme(legend.position = "none") + scale_color_manual(values = colorscharmaip)

plot_grid(a,b)

LDA.hyp %>% distinct(treatment, predictedstage) %>%
  group_by(treatment) %>%
  summarize(predictedstages = str_c(predictedstage , collapse = ", "))

LDA.hyp %>% 
  group_by(treatment,predictedstage) %>%
  summarize(n = n())

LDA.pit %>% distinct(treatment, predictedstage) %>%
  group_by(treatment) %>%
  summarize(predictedstages = str_c(predictedstage , collapse = ", "))

LDA.pit %>% 
  group_by(treatment,predictedstage) %>%
  summarize(n = n())

LDA.gon %>% distinct(treatment, predictedstage) %>%
  group_by(treatment) %>%
  summarize(predictedstages = str_c(predictedstage , collapse = ", "))

LDA.gon %>% 
  group_by(treatment,predictedstage) %>%
  summarize(n = n())

```
